---
title: "Estimates extraction for 'Using meta-analysis for evidence synthesis: The case of incomplete neutralization in German'"
author: 
  - name: "Bruno Nicenboim, Timo B. Roettger, and Shravan Vasishth"
date: 
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: true
    css: styles.css
bibliography: edaps.bib
---

# Estimate extraction

We provide here the code showing how we extracted the estimates that went into our Bayesian meta-analysis from the summaries provided in the publications or from the data obtained from the authors.

```{r setup, cache=FALSE,include=FALSE}
set.seed(123)
# global chunk options
library(knitr)
library(rstudioapi)
opts_chunk$set(cache=TRUE, fig.path='fig_rmd/', cache.path='cache_estimates/', fig.align='center', tidy=FALSE) 
opts_knit$set(self.contained=FALSE)

current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_working_dir)

```

## Setup

The following packages should be loaded:

```{r libraries, message=F}
library(knitr)
library(dplyr)
library(tidyr)
library(purrr)
library(magrittr)
library(rstan)
library(stringr)
library(brms)
library(tibble)

sessionInfo() 
```

It is recommended to run Stan [@Stan2017] code with 3 cores (or more) and at least 2000 iterations. For our models, we used the wrapper package brms [@brms].

```{r options}
# Uses the computer cores available
options(brms.backend = "cmdstanr")
options(mc.cores = parallel::detectCores())
# The iterations can be reduced *only* for testing purposes:
iter <- 2000
chains <- 4
# options to increase efficiency of brms models (optional)

```

## Data extraction

We extracted data regarding the preceding vowel duration in ms, and we summarized the posterior of the difference between voiceless (e.g. *t* in *Rat*) and devoiced consonants (e.g., *d* in *Rad*) by storing the mean estimate (`Estimate`), $\hat\beta$, the 95% credible interval (`2.5%ile` and `97.5%ile`), the probability that the difference in vowel duration would be positive (`Pbeta`), and the SD of the posterior distribution (`Est.Error`). We summarize the data extracted at the end of this report.


```{r extraction}
# To extract the relevant measure from the data files:
PVD <- "preceding vowel duration in ms"
# Summaries will be saved here:
results <- list()
```

We modified the labels of the datasets so that for every study, the following variables have the same meaning:

* `dur` is the *dependent variable* vowel duration.
* `cvoicing` is the *predictor* that indicates whether the last stop is (de)voiced $.5$ or $-.5$ voiceless.
* Group-level (random) effects (not all of them appear in every study):
	- `pair` indicates whether the observation belongs to the same "pair", e.g., *Rad* and *Rat* belong to the same pair, and *Tap* and *Tab* to another pair.
	- `subject`.
	- `elicitation_method` depends on the study, and they include read list, repeat, read contrastive, etc. We coded $1$ for reading methods, and $-1$ for non reading methods.

### Data extracted from complete datasets

We used Bayesian linear mixed models with the maximal random effects structure and weakly informative regularizing priors for the complete datasets that were available to us: @Fuchs_2005, @Grawunder_2014}, Experiment~1 and 2 of @Roettger_etal_2014  and Experiment 1 and 2 of @Baer-Henney_unpublished.

We fitted all the models with *brms*, the meaning of the priors of the models can be extracted using  `get_prior`, and the Stan code can be seen using `make_stancode`.

The data is available in our [osf repository](https://osf.io/4c25h/), and needs to be saved in a `Data` folder where this analysis is run.

##### @Fuchs_2005

```{r F05, message=F}
df_F05 <- read.table("../Data/Fuchs_2005_complete.tsv", sep="\t", header = T)  %>% mutate(cvoicing = if_else(voicing == "voiced", .5, -.5))

head(df_F05)

# Linear mixed model:
fit_F05 <- brm(dur ~ cvoicing + (cvoicing | pair) +(cvoicing | subject), 
            data = df_F05,
            prior = c(prior(normal(0,300), class = Intercept),
                prior(normal(0,100), class = b, coef = cvoicing),
                prior(normal(0,100), class = sd),
                prior(normal(0,100), class = sigma),
                prior(lkj(2), class = cor)),
            iter = iter,chains = chains,
            control = list(adapt_delta = .9999, max_treedepth = 15))

fit_F05

results[["F05"]]  <- fixef(fit_F05)["cvoicing", ]
results[["F05"]]["intercept"]  <- fixef(fit_F05)["Intercept", 1]
results[["F05"]]["prop"]  <- fixef(fit_F05)["cvoicing", 1] / fixef(fit_F05)["Intercept", 1] 
results[["F05"]]["Pbeta"] <- mean(posterior_samples(fit_F05, pars = "b_cvoicing") > 0)
```

#### @Grawunder_2014

```{r G14, message=F}
df_G14 <- read.table("../Data/Grawunder_2014.tsv", sep = "\t", header = T) %>%
         mutate(cvoicing = if_else(voicing == "voiced",.5,-.5))

fit_G14 <- brm(dur ~ cvoicing + (cvoicing|pair) + 
                (cvoicing|subject), 
                data = df_G14,
                prior = c(prior(normal(0,300), class = Intercept),
                prior(normal(0,100), class = b, coef = cvoicing),
                prior(normal(0,100), class = sd),
                prior(normal(0,100), class = sigma),
                prior(lkj(2), class = cor)),
            iter = iter, chains = chains,
            control = list(adapt_delta = .9999, max_treedepth = 15))

fit_G14

results[["G14"]]  <- fixef(fit_G14)["cvoicing",]
results[["G14"]]["intercept"]  <- fixef(fit_G14)["Intercept", 1]
results[["G14"]]["prop"]  <- fixef(fit_G14)["cvoicing", 1] / fixef(fit_G14)["Intercept", 1] 
results[["G14"]]["Pbeta"] <- mean(posterior_samples(fit_G14, 
                                            pars = "b_cvoicing") > 0)
```

#### Experiment 1 of @Roettger_etal_2014

```{r Rea14_1, message=F}
df_Rea14_1 <- read.table("../Data/Roettger_et_al_2014_E1_production.tsv", 
                sep="\t", header=T,comment.char="",   quote = "" ) %>% 
                filter(usable == "usable")  %>% 
                mutate(cvoicing = if_else(voicing == "voiced", .5, -.5)) %>% 
                mutate(pair = tolower(str_sub(str_trim(item), 1, -2))) %>% 
                mutate(dur = vowel_dur)


head(df_Rea14_1)

# Linear mixed model:
fit_Rea14_1 <- brm(dur ~ cvoicing + (cvoicing | pair) + 
                (cvoicing | subject), 
                    data = df_Rea14_1,
                prior = c(prior(normal(0,300), class = Intercept),
                    prior(normal(0,100), class = b, coef = cvoicing),
                    prior(normal(0,100), class = sd),
                    prior(normal(0,100), class = sigma),
                    prior(lkj(2), class = cor)), 
                iter = iter, chains = chains,
                control = list(adapt_delta = .999, max_treedepth = 15))

fit_Rea14_1

results[["Rea14_1"]]  <- fixef(fit_Rea14_1)["cvoicing",]
results[["Rea14_1"]]["intercept"]  <- fixef(fit_Rea14_1)["Intercept", 1]
results[["Rea14_1"]]["prop"]  <- fixef(fit_Rea14_1)["cvoicing", 1] / fixef(fit_Rea14_1)["Intercept", 1] 
results[["Rea14_1"]]["Pbeta"] <- mean(posterior_samples(fit_Rea14_1, 
                                            pars = "b_cvoicing") > 0)
```


#### Experiment 2 of @Roettger_etal_2014

```{r Rea14_2, message=F}
df_Rea14_2 <- read.table("../Data/Roettger_et_al_2014_E2_production.tsv", 
                sep= "\t", header = T,comment.char = "", quote = "" ) %>% 
               filter(usable == "usable", comment == "Normal")  %>% 
               mutate(cvoicing = if_else(voicing == "voiced", .5, -.5)) %>% 
               mutate(pair = tolower(str_sub(str_trim(item),1,-2))) %>% 
               mutate(dur = vowel_dur)

 
head(df_Rea14_2)

# Linear mixed model:
fit_Rea14_2 <- brm(dur ~ cvoicing + (cvoicing|pair) + 
                (cvoicing|subject), 
                    data = df_Rea14_2,
                prior = c(prior(normal(0,300), class = Intercept),
                    prior(normal(0,100), class = b, coef = cvoicing),
                    prior(normal(0,100), class = sd),
                    prior(normal(0,100), class = sigma),
                    prior(lkj(2), class = cor)),
                iter = iter, chains = chains,
                control = list(adapt_delta = .999, max_treedepth = 15))

fit_Rea14_2

results[["Rea14_2"]]  <- fixef(fit_Rea14_2)["cvoicing",]
results[["Rea14_2"]]["intercept"]  <- fixef(fit_Rea14_2)["Intercept", 1]
results[["Rea14_2"]]["prop"]  <- fixef(fit_Rea14_2)["cvoicing", 1] / fixef(fit_Rea14_2)["Intercept", 1] 
results[["Rea14_2"]]["Pbeta"] <- mean(posterior_samples(fit_Rea14_2, 
                                            pars = "b_cvoicing") > 0)
```

#### Experiment 1 of @Baer-Henney_unpublished

```{r B_1, message=F}
df_B_1 <- read.table("../Data/Baer-Henney_unpublished.tsv", sep="\t", header=T) %>%
             mutate(cvoicing = if_else(voicing == "voiced", .5, -.5))

head(df_B_1)

fit_B_1 <- brm(dur ~ cvoicing + (cvoicing|pair) + 
                (cvoicing|subject), 
                data = df_B_1,
            prior = c(prior(normal(0,300), class = Intercept),
                    prior(normal(0,100), class = b, coef = cvoicing),
                    prior(normal(0,100), class = sd),
                    prior(normal(0,100), class = sigma),
                    prior(lkj(2), class = cor)),
            iter = iter, chains = chains,
            control = list(adapt_delta = .999, max_treedepth = 15))

fit_B_1

results[["B_1"]]  <- fixef(fit_B_1)["cvoicing",]
results[["B_1"]]["intercept"]  <- fixef(fit_B_1)["Intercept", 1]
results[["B_1"]]["prop"]  <- fixef(fit_B_1)["cvoicing", 1] / fixef(fit_B_1)["Intercept", 1] 
results[["B_1"]]["Pbeta"] <- mean(posterior_samples(fit_B_1, 
                                            pars = "b_cvoicing") > 0)
```

#### Experiment 2 of @Baer-Henney_unpublished

```{r B2, message = F}

df_B_2 <- read.table('../Data/Baer-Henney_unpublished_II.tsv', sep = "\t", 
            header = T ) %>%
           mutate(cvoicing = if_else(voicing == "voiced", .5, -.5),
            dur = dur*1000)

head(df_B_2)

fit_B_2 <- brm(dur ~ cvoicing + (cvoicing|pair) + 
                (cvoicing|subject), 
                data = df_B_2,
            prior = c(prior(normal(0,300), class = Intercept),
                prior(normal(0,100), class = b, coef = cvoicing),
                prior(normal(0,100), class = sd),
                prior(normal(0,100), class = sigma),
                prior(lkj(2), class = cor)),
            iter = iter, chains = chains,
            control = list(adapt_delta = .999, max_treedepth = 15))

fit_B_2

results[["B_2"]]  <- fixef(fit_B_2)["cvoicing",]
results[["B_2"]]["intercept"]  <- fixef(fit_B_2)["Intercept", 1]
results[["B_2"]]["prop"]  <- fixef(fit_B_2)["cvoicing", 1] / fixef(fit_B_1)["Intercept", 1] 
results[["B_2"]]["Pbeta"] <- mean(posterior_samples(fit_B_2, 
                                            pars = "b_cvoicing") > 0)
```

### Data summarized at some level with standard deviations

This was the case when studies presented data that were already summarized at some
level (some combination of by-items, by-participants and/or by-repetitions) and presented SDs (and the raw data was not available): @Mitleb_1981, @Fourakis_Iverson_1984,  @Port_ODell_1985, @Port_Crawford_1989, @Greisbach_2001, and @Piroth_Janker_2004.

In these cases we fitted measurement error mixed models and we computed the SE of each data cell (i.e., the provided mean of a group of observations) by dividing the SD (`sd` in the datasets) provided in the paper by the square root of the number of observation per cell (`N` in the datasets). 

#### @Mitleb_1981

```{r M81, message=F}
df_M81 <- read.table("../Data/Mitleb_1981.tsv", sep="\t", header=T) %>%
          filter(acoustic_cue == PVD)  %>%
          mutate(cvoicing = if_else(voicing == "voiced", .5, -.5),
                obs =1:n())

head(df_M81)

## Sigma of the model (non-centered parametrization) is achieved with (1|obs)
fit_M81_me <- brm(dur | se(sd/sqrt(N)) ~ cvoicing + (1|obs) + (1|pair), 
                            data=df_M81,
                            prior= c(prior(normal(0,300), class = Intercept),
                                     prior(normal(0,100), class = b, coef = cvoicing),
                                     prior(normal(0,100), class = sd)),
                            control=list(adapt_delta=.999, max_treedepth=12),
                            iter=iter)

fit_M81_me

results[["M81"]]  <- fixef(fit_M81_me)["cvoicing",]
results[["M81"]]["intercept"]  <- fixef(fit_M81_me)["Intercept", 1]
results[["M81"]]["prop"]  <- fixef(fit_M81_me)["cvoicing", 1] / fixef(fit_M81_me)["Intercept", 1] 
results[["M81"]]["Pbeta"] <- mean(posterior_samples(fit_M81_me, 
                                            pars = "b_cvoicing") > 0)
```
#### @Fourakis_Iverson_1984

```{r FI84, message=F}

df_FI84 <- read.table("../Data/Fourakis_Iverson_1984.tsv", sep="\t", header=T) %>% 
            filter(acoustic_cue == PVD)  %>% 
            mutate(cvoicing=if_else(voicing == "voiced",.5,-.5),
                    obs= 1:n(),
                    elic = ifelse(study == 2, 1,-1)) #study 2 was reading
df_FI84 <- df_FI84[complete.cases(df_FI84),]


head(df_FI84)

fit_FI84_me <- brm(dur | se(sd/sqrt(N)) ~ cvoicing * elic + (1|obs) + 
                    (cvoicing|pair) + (cvoicing|subject), data = df_FI84,
                    prior = c(prior(normal(0,300), class = Intercept),
                        prior(normal(0,100), class = b),
                        prior(normal(0,100), class = sd),
                        prior(lkj(2), class = cor)),
                    control=list(adapt_delta=.9999, max_treedepth=16), iter = iter)

fit_FI84_me

results[["FI84"]]  <- fixef(fit_FI84_me)["cvoicing",]
results[["FI84"]]["intercept"]  <- fixef(fit_FI84_me)["Intercept", 1]
results[["FI84"]]["prop"]  <- fixef(fit_FI84_me)["cvoicing", 1] / fixef(fit_FI84_me)["Intercept", 1] 
results[["FI84"]]["Pbeta"] <- mean(posterior_samples(fit_FI84_me, 
                                            pars = "b_cvoicing") > 0)
```

#### @Port_ODell_1985

```{r PO85, message=F}
df_PO85 <- read.table("../Data/Port_ODell_1985.tsv", sep="\t", header=T) %>% 
           filter(acoustic_cue == PVD)  %>% 
           mutate(cvoicing = if_else(voicing == "voiced", .5, -.5),
                  dur = dur*10,
# Vowel duration was presented in cs rather than ms:
                  sd = sd*10,
                  obs = 1:n())

head(df_PO85)

fit_PO85_me <- brm(dur | se(sd/sqrt(N)) ~ cvoicing + (1|obs) + 
                    (cvoicing|pair), data=df_PO85,
                prior = c(prior(normal(0,300), class = Intercept),
                         prior(normal(0,100), class = b, coef = cvoicing),
                         prior(normal(0,100), class = sd),
                         prior(lkj(2), class = cor)),
                control=list(adapt_delta=.999, max_treedepth=15), iter = iter)

fit_PO85_me

results[["PO85"]]  <- fixef(fit_PO85_me)["cvoicing",]
results[["PO85"]]["intercept"]  <- fixef(fit_PO85_me)["Intercept", 1]
results[["PO85"]]["prop"]  <- fixef(fit_PO85_me)["cvoicing", 1] / fixef(fit_PO85_me)["Intercept", 1] 
results[["PO85"]]["Pbeta"] <- mean(posterior_samples(fit_PO85_me, 
                                            pars = "b_cvoicing") > 0)
```

#### @Port_Crawford_1989

```{r PC89, message=F}
df_PC89 <- read.table("../Data/Port_Crawford_1989.tsv", sep="\t", header=T) %>%
           filter(acoustic_cue == PVD)  %>%
           mutate(cvoicing=if_else(voicing=="voiced",.5,-.5),
                  obs = 1:n(),
                  elic = ifelse(elicitation_method %in% 
                            c("read contrastive","read list"), 1, -1) )
df_PC89 <- df_PC89[complete.cases(df_PC89),]

head(df_PC89)

fit_PC89_me <- brm(dur | se(sd/sqrt(N)) ~ cvoicing * elic 
                    + (1 | obs) + (cvoicing * elic | pair), data = df_PC89,
                    prior= c(prior(normal(0,300), class = Intercept),
                             prior(normal(0,100), class = b),
                             prior(normal(0,100), class = sd),
                             prior(lkj(2), class = cor)),
                control=list(adapt_delta=.999, max_treedepth=15), iter=iter)

fit_PC89_me

results[["PC89"]]  <- fixef(fit_PC89_me)["cvoicing",]
results[["PC89"]]["intercept"]  <- fixef(fit_PC89_me)["Intercept", 1]
results[["PC89"]]["prop"]  <- fixef(fit_PC89_me)["cvoicing", 1] / fixef(fit_PC89_me)["Intercept", 1] 
results[["PC89"]]["Pbeta"] <- mean(posterior_samples(fit_PC89_me, 
                                            pars = "b_cvoicing") > 0)

```
#### @Greisbach_2001

```{r G01, message=F}
df_G01 <- read.table("../Data/Greisbach_2001.tsv", sep="\t", header=T) %>%
          filter(acoustic_cue == PVD)  %>%
          mutate(cvoicing = if_else(voicing == "voiced", .5, -.5),
                 obs = 1:n())


fit_G01_me <- brm(dur | se(sd/sqrt(N)) ~ cvoicing + (1 | obs) + 
                                            (cvoicing | pair), data = df_G01,
                  prior = c(prior(normal(0,300), class = Intercept),
                           prior(normal(0,100), class = b, coef = cvoicing),
                           prior(normal(0,100), class = sd),
                           prior(lkj(2), class = cor)),
                  control=list(adapt_delta=.999, max_treedepth=15), iter = iter)

fit_G01_me

results[["G01"]]  <- fixef(fit_G01_me)["cvoicing",]
results[["G01"]]["intercept"]  <- fixef(fit_G01_me)["Intercept", 1]
results[["G01"]]["prop"]  <- fixef(fit_G01_me)["cvoicing", 1] / fixef(fit_G01_me)["Intercept", 1] 
results[["G01"]]["Pbeta"] <- mean(posterior_samples(fit_G01_me, 
                                            pars = "b_cvoicing") > 0)
```


#### @Piroth_Janker_2004

```{r PJ04, message=F}

df_PJ04 <- read.table("../Data/Piroth_Janker_2004.tsv", sep="\t", header=T) %>%
           filter(acoustic_cue == PVD)  %>%
           mutate(cvoicing = if_else(voicing == "voiced", .5, -.5),
                   obs = 1:n())
df_PJ04 <- df_PJ04[complete.cases(df_PJ04),]


head(df_PJ04)

fit_PJ04_me <- brm(dur | se(sd/sqrt(N)) ~ cvoicing + (1|obs) + 
                    (cvoicing|subject), data = df_PJ04,
                prior = c(prior(normal(0,300), class = Intercept),
                         prior(normal(0,100), class = b, coef = cvoicing),
                         prior(normal(0,100), class = sd),
                         prior(lkj(2), class = cor)),
                control=list(adapt_delta=.999, max_treedepth=15), iter = iter)

fit_PJ04_me

results[["PJ04"]]  <- fixef(fit_PJ04_me)["cvoicing",]
results[["PJ04"]]["intercept"]  <- fixef(fit_PJ04_me)["Intercept", 1]
results[["PJ04"]]["prop"]  <- fixef(fit_PJ04_me)["cvoicing", 1] / fixef(fit_PJ04_me)["Intercept", 1] 
results[["PJ04"]]["Pbeta"] <- mean(posterior_samples(fit_PJ04_me, 
                                            pars = "b_cvoicing") > 0)
```


### Data summarized without standard deviations or any measure of dispersion

This was the case when a study presented data that were already summarized at some
level but no measure of dispersion was provided: @Charles-Luce_1985

In this case, we fit a linear  model that ignores the measurement error.

#### @Charles-Luce_1985

```{r C85, message=F}
df_C85 <- read.table("../Data/Charles-Luce_1985.tsv", sep="\t", 
                        header=T,comment.char="") %>% 
            filter(acoustic_cue == PVD)  %>% 
            mutate(cvoicing = if_else(voicing == "voiced", .5, -.5))

head(df_C85)

fit_C85 <- brm(dur ~  cvoicing, data = df_C85,
                prior= c(prior(normal(0,300), class = Intercept),
                         prior(normal(0,100), class = b, coef = cvoicing),
                         prior(normal(0,100), class = sigma)),iter=iter,
                control=list(adapt_delta=.999, max_treedepth=15))

fit_C85

results[["C85"]]  <- fixef(fit_C85)["cvoicing",]
results[["C85"]]["intercept"]  <- fixef(fit_C85)["Intercept", 1]
results[["C85"]]["prop"]  <- fixef(fit_C85)["cvoicing", 1] / fixef(fit_C85)["Intercept", 1] 
results[["C85"]]["Pbeta"] <- mean(posterior_samples(fit_C85, 
                                            pars = "b_cvoicing") > 0)
```

### No summary of the data or data available

In this case only mean, and F-values were provided. We calculated the SE, and we approximated the confidence interval: @Smith_etal_2009 

The absolute avg. vowel duration is visually extracted from figure 2. 159ms for voiced and 146ms for voiceless.

#### @Smith_etal_2009

```{r, message=F}
Mean <- 13 
F <- 4.96
(SE <- Mean/sqrt(F))
(lowerCI <- Mean - 2 * SE)
(upperCI <- Mean + 2 * SE)
results[["Sea09"]]  <- c(Mean, SE, lowerCI, upperCI, 146, 0.0890411, NA)
```

## Summary of estimates extracted

```{r, message=F, results = 'asis'}
df_results <- plyr::ldply(results, .id = "short_name") 
kable(df_results) 
write.csv(df_results, "Estimates.csv") 
```

# References