---
title: "Adjusted Script from Nicenboim et al. (2018) for ICPHS"
author: 
  - name: "Timo B. Roettger"
date: 
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: true
    css: styles.css
bibliography: edaps.bib
---

# Meta-analysis

```{r setup, cache=FALSE,include=FALSE}

# read in libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(viridis, tidyverse, here, knitr, kableExtra, brms, purrr)

# global seed
set.seed(123)
# global chunk options
opts_chunk$set(cache=TRUE, fig.path='fig_rmd/', cache.path='cache_estimates/', fig.align='center') 
opts_knit$set(self.contained=FALSE) 


# function to ignoring the setting of the relative path below when knitting
# run_if_not_knitting <- function(expr) {
#   if (!isTRUE(getOption("knitr.in.progress"))) {
#     eval(expr)
#   }
# }
# 
# # set the current working directory to the one where this file lives
# run_if_not_knitting(current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path))
# run_if_not_knitting(setwd(current_working_dir))

```


It is recommended to run Stan [@Stan2017] code with 3 cores (or more) and at least 2000 iterations. For our models, we used the wrapper package brms [@brms].

```{r options}
# Uses the computer cores available
options(mc.cores = parallel::detectCores())
options(brms.backend = "cmdstanr")

# The iterations can be reduced *only* for testing purposes:
iter <- 2000
chains <- 4
```

## Meta analysis

The estimates can be generated by the script: estimate_extraction.Rmd

```{r estimates}
all_1 <- read.csv(here::here("Derived_Data","Estimates.csv")) |> 
  mutate(prop.error = Est.Error / intercept)

head(all_1) 
```

As opposed to the original meta-analysis, we estimate the proportion of vowel duration difference relative to the voiced vowel.

```{r brmsma}


# priors
mypriors <- c(brms::prior(normal(0,0.3), class = Intercept),
              brms::prior(normal(0,0.3), class = sd))

fit_ma_all <- brms::brm(prop | se(`prop.error`) ~ (1|short_name) , data = all_1,
                  prior = mypriors,      
                  control = list(adapt_delta = .999, 
                               max_treedepth = 12),
                  iter = iter,
                  file_refit = "on_change",
                  file = here::here("Models", "meta_analysis")
    )

fit_ma_all 

```

What is the average proportion of measurement error

```{r}

all_1_props <- all_1 |> 
  mutate(prop_1ms = (1/intercept)*100,
         prop_2ms = (2/intercept)*100,
         prop_3ms = (3/intercept)*100,
         prop_4ms = (4/intercept)*100,
         prop_5ms = (5/intercept)*100
         ) |> 
  summarise(prop_1ms_avg = mean(prop_1ms),
            prop_2ms_avg = mean(prop_2ms),
            prop_3ms_avg = mean(prop_3ms),
            prop_4ms_avg = mean(prop_4ms),
            prop_5ms_avg = mean(prop_5ms)
            )

```

